---
title: "Introduction to BubbleTime"
author: "Matthew H. Grinnell"
date: "`r Sys.Date()`"
output:
      rmarkdown::html_vignette:
        fig_caption: yes
        toc: true
        number_sections: true
bibliography: "../inst/PACKAGES.bib"
csl: "../inst/cjfas.csl"
link-citations: true
vignette: >
  %\VignetteIndexEntry{Introduction to BubbleTime}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# Set knitr options
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5.75
)
# Set knitr options
options(knitr.kable.NA = "")
```

```{r citation, include = FALSE}
# Write citation file
knitr::write_bib(x = "BubbleTime", file = here::here("inst", "PACKAGES.bib"))
```

This vignette shows how to calculate dive time.
Source code is on the
[GitHub repository](https://github.com/grinnellm/BubbleTime).
Load the `BubbleTime` package [@R-BubbleTime] and
a few other packages to wrangle data and make tables and figures.

```{r libraries, message = FALSE, warning = FALSE}
# Install the package if required
# install.packages("remotes")
# remotes::install_github(repo = "grinnellm/BubbleTime")
# Load the package
library(BubbleTime)
# Suppress dplyr messages re summarise
options(dplyr.summarise.inform = FALSE)
```

# Dive time calculations

This section shows two ways to calculate dive times and view results.
The calculations are the same in both cases;
the difference is that the second way uses a wrapper
to run all the functions in one line.

## The long way

First, load dive times from a CSV file.
In this example I am using the provided data `dives`.
Note the column names and format in each column.
Your data should look like this when you load it into **R**.
See the help file for `dives` for more information.

```{r load_dives}
# Load dive times from file
dives_raw <- load_dives(
  file_path = system.file("extdata", "dives.csv", package = "BubbleTime")
)
# Print to screen
dives_raw
```

Next, calculate the duration of each dive in minutes.

```{r calc_raw_mins}
# Calculate raw dive time
raw_mins <- calc_raw_mins(dat = dives_raw)
# Print to screen
raw_mins
```

Take a look at raw dive times for each diver to check whether
the data is correct. 

```{r plot_daily_mins}
# Plot daily dive time
plot_daily_mins(dat = raw_mins)
```

Finally, calculate daily dive time in hours.

```{r calc_daily_hrs}
# Calculate adjusted daily dive time in hours
daily_hrs <- calc_daily_hrs(dat = raw_mins)
# Print to screen
daily_hrs
```

Again, look at adjusted daily dive times for each diver to check whether
the data is correct. 

```{r plot_cumulative_hrs}
# Plot cumulative dive time
plot_cumulative_hrs(dat = daily_hrs)
```

## The easy way

Alternately, take the easy route with some help from your `dive_buddy()`.
This is just a wrapper around the functions above.

```{r dive_buddy_code}
# Run helper function with included data file
dive_buddy(
  file_path = system.file("extdata", "dives.csv", package = "BubbleTime")
)
```

You can also run `dive_buddy()` interactively
(i.e., to find your input file)
and save the output.

```{r dive_buddy_manual, eval = FALSE}
# Run helper function interactively and save output (not run in this example)
dive_buddy(save_plot = TRUE, save_csv = TRUE)
```

# References
